The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2017-08-31 JStatSoft_forestinventory_master-concordance.tex
update Andreas 31.08.2017

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 123 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
139 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
1 77 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
175 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
192}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
194 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1483:%
206 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1492:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 21 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1698:%
216 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1714:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1972:%
233 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1976:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
238 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 104 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 112:%
1 94 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 207:%
120 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 217:%
1 75 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 293:%
131 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 305:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 836:%
144}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 837:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1356:%
146 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1368:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1433:%
159 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1443:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1740:%
170 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1750:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2008:%
181 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2017:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2022:%
191 11 1}

### 2017-08-31 JStatSoft_forestinventory_master.rnw
update Andreas 31.08.2017

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Andreas Hill\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{andreas.hill@usys.ethz.ch}\\
  Telephone: +41/44/632 32 36\\
  URL: \url{http://www.lue.ethz.ch/people/hilla}

@@

%% for those who use Sweave please include the following line (with % symbols):

\usepackage{Sweave}

@@

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\LF}{\ensuremath{\lambda(F)}}
\newcommand{\LFC}{\ensuremath{\lambda^2(F)}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\PR}{\mathbb{P}}
\newcommand{\var}{\mathbb{V}}
% \newcommand{\COV}{\mathbb{COV}}
\newcommand{\MAV}{\mathbb{MAV}}
\newcommand{\MRAV}{\mathbb{MRAV}}
\newcommand{\POP}{\mathcal{P}}
\newcommand{\SAMP}{\mathcal{S}}
\newcommand{\RE}{\mathbb{RE}}
\newcommand{\PLAN}{\Re^2}
\newcommand{\SUR}{\mathbb{S}}
\newcommand{\ING}{\mathbb{I}}
\newcommand{\DEP}{\mathbb{D}}

%\setlength{\parindent}{1em}
%\setlength{\mathindent}{0pt}

@@

% NOTES:
% --> take parts of the old introduction version of the diploma-thesis!

% NOTES:
% 1st section: - general introduction to forest inventories --> model-'supported' inventories --> 
%                model-dependent vs. design-based
% 2nd section: + (more R-specific): what's already there? --> mention sampling / inventory packages
%                 and seperate them from our package (focus, applied areas, approach)
%              + important reasoning for our package: we are the first that implemented and provide 
%                a large range of design-based estimators
% 3rd section: what is the motivation of the package: - to make the broad range of Mandallaz-estimators
%              available to a large user community, to facilitate their application in science as well as 
%              operational forest management (why is this important? --> because the methods have demonstrate
%              the potential of considerably increasing the estimation precision)
% 4th section: whats the purpose of the paper? --> the objective of the paper is to 
%             a) establish the link between the mathematical description of the estimators (given in Mandallaz ...) 
%                and their implementation in our package
%             b) illustrate the application of the various estimators by the respsective functions in our package 
%                to real-world inventory scenarios
%             c) highlight 'special cases' , i.e. rare inventory scenarios, and demonstrate how the package-functions
%                deal with such situations (error-checking functions and data-adjustments)

% Idee: Reine Theory und Applikation NICHT stark trennen --> Jeweils für jeden Schätzer die nötigsten Formeln geben 
%       (Punktschätzung und g-weight und external variance, wenn nötig noch zusätzliche Infos wie Dummy Variable)
%      WICHITG: Davor muss das generelle Konzept kommen (Entscheidungsbaum und konzeptionelle Erklärung (z.B.: was heisst
%               "exhaustive" usw.) --> auf der Graphik dann ansetzen
%
% evtl auch noch die Art von Graphik für den Aufbau einer 2- und 3-phaisgen Inventurs (wie in den Präsentationen)

%------------------------------------------------------------------------------------------------%
% --------------------------- Structure of the Package ----------------------------------------- %

@@

% NOTES:
% - Pyramide / decision tree Graphik
% - Confidence Interval calculation
% - Analysis of estimations results. a) Visualization functions; b) Gain-analysis
%
% --> explain each point briefly without to much (mathematical) detail

@@

% NOTES:
% 1st subsection: Global 2 and 3-phase estimators for simple and cluster sampling
% 2nd subsection: 2- and 3-phase Small Area estiamtors for simple and cluster sampling
%

@@

% NOTES:
% - demonstrate calculation of confidence intervals (give the formulas)

@@

% NOTES:
% - Error handling and adjustments: Sampling designs not nested (always give background and then illustrate handling in package)
% - Error warning: unbalanced designs and singularities due to missing factor levels in s2-sample (2-phase)
%                   or s2 or/and s1 sample in 3-phase case
% - Error warning: cluster not completely within one small area (--> g-weight variance estimator) +  recommendation

% \SweaveInput{rnw/.Rnw}

@@

%------------------------------------------------------------------------------------------------%

>>>>

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Andreas Hill\\
  Department of Environmental Systems Science\\
  Chair of Landuse Engineering\\
  ETH Z\"urich\\
  Universit\"atstrasse 16\\
  8092 Z\"urich, Switzerland\\
  E-mail: \email{andreas.hill@usys.ethz.ch}\\
  Telephone: +41/44/632 32 36\\

@@

%% for those who use Sweave please include the following line (with % symbols):
\usepackage{Sweave}

@@

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\var}{\mathbb{V}}

@@

% NOTES:
%
%
%
%

%------------------------------------------------------------------------------------------------%
% --------------------------- Methods and structure of the Package ----------------------------- %

@@

% NOTES:
% - mention difference between finite and infinite population apporach ?
%
%
%

@@

% NOTES:
%
%
%
%

@@

% NOTES:
%
%

@@

% NOTES:
%
%
%

@@

% NOTES:
% 
%
%

%------------------------------------------------------------------------------------------------%
% ---------------------------------------- Acknowledgements ------------------------------------ %

### 2017-08-31 rnw/Analysis and Visualization.Rnw
update Andreas 31.08.2017

The package \pkg{forestinventory} also provides methods for \textit{analysing} and \textit{visualizing} the estimation results created by the \code{onephase}, \code{twophase} and \code{threephase} functions.

@@

If we want to compare the results and performance of different estimation methods and estimators for a given global or small area inventory data set, the function to start with is \code{estTable()}. This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} dataset:

@@

The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}. If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}. Note however that the package \pkg{forestinventory} provides several \code{S3} generic functions that particularly respond to the class \code{"esttable"}. Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.

The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package. However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or threephase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied. Per default, also the confidence intervals are added.

@@

The package \pkg{forestinventory} also provides a \code{S3} generic plotting function based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results. Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function. The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.

If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case for the default settings of \code{estTable()}:

@@

The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic plot function and setting the argument \code{yvar = "estimate"}:

@@

If we want the estimation errors to be plotted, we simply leave the \code{yvar} argument to its default (\code{"error"}):

@@

Note that the \code{plot}-function call can arbitrarily be extended by additional \pkg{ggplot2} functions and specifications.

>>>>

The package \pkg{forestinventory} also provides methods for \textit{analyzing} and \textit{visualizing} the estimation results created by the \code{onephase()}, \code{twophase()} and \code{threephase()} functions.

@@

If we want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, the function to start with is \code{estTable()}. This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} data set:

@@

The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}. If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}. Note however that the package \pkg{forestinventory} provides several \code{S3} generic methods that particularly respond to the class \code{"esttable"}. Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.

The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package. However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied. Per default, also the confidence intervals are added.

@@

The package \pkg{forestinventory} also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results. Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function. The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.

If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case using the default setting of \code{estTable()}:

@@

The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic \code{plot} method and setting the argument \code{yvar = "estimate"}:

@@

If we want the estimation errors to be plotted, we simply set the \code{yvar} argument \code{"error"} (default):

@@

Note that the \code{plot()}-function call can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.

### 2017-08-31 rnw/Introduction.Rnw
update Andreas 31.08.2017

In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods. They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level. Whereas the most straightforward approach of collecting the required information would be a full census of all trees within the forest area of interest, this remains only feasible for spatially small areas and is even then scarcely done in practice due to time- and cost restrictions. For this reason, forest inventories gather their information of interest by means of statistical sampling methods, i.e. information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}. This sample is then used to provide estimations of a target variable for a given spatial unit, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}. The information of terrestrial samples is assumed to be very precise, and increasing the precision of the estimates could primarily be achieved by increasing the terrestrial sample size. However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited. Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par

@@

% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
Despite these promising developments, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software. One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach. However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing. Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The package comprises two and cover 32 inventory scenarios in total. The estimators can be used for stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The particular objective of this article is a) to establish the link between the mathematical description of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e. rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).

>>>>

In many countries, forest inventories have become an integral instrument to judge the current state as well as the development of forests within reoccurring time periods. They provide quantitative information when it comes to define management actions and to adapt forest management strategies according to guidelines on national- and international level. As a full census of all trees within the target forest area only remains feasible for spatially small areas, forest inventories commonly gather their information of interest by means of statistical sampling methods, i.e. information is gathered only at discrete sample locations (\textit{sample plots}) in the forest in the framework of a \textit{terrestrial inventory}. This sample is then used to provide a \textit{point estimate} of the true, albeit unknown population parameter of the forest area, and there is a broad range of concepts and methods regarding the choice of the sample design and respective estimators \citep{gregoire2007, kohl2006, schreuder1993, mandallaz2008}. The information of terrestrial samples is assumed to be very precise, and increasing the precision of the point estimate - characterized by its \textit{variance} - could primarily be achieved by increasing the terrestrial sample size. However, since collecting the terrestrial information is very time consuming and expensive, the number of terrestrial samples is usually limited. Whereas in national inventories the terrestrial sample size is still sufficient to provide high estimation accuracies on the national scale, small sample sizes on smaller spatial scales such as forest management units often cause high estimation errors that hamper using the inventory information.\par

@@

Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle such situations.

### 2017-08-31 rnw/Special Cases and Scenarios.Rnw
update Andreas 31.08.2017

To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} dataset into 3 classes:

@@

As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample location, i.e. sample unit, does no longer consist of one, but multiple spatially agglomerated sample points. Whereas the maximal possible number of sample points per cluster $M$ is defined by the user, the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random. Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix). The package functions allow to consider cluster-structures by storing aspecifying the argument \code{cluster}

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured.

@@

Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that the point estimates are close and more important, the external as well as the design-based variances only differ marginally. This can be taken as a hint that the violation of the zero mean residual property can expected to be negligible in this case. We can also check the calculated mean of the residuals for the small area by

@@

As explained in section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other. The correct nesting thereby concerns the spatial arrangement of the sample phases (figure \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location. For the latter, the package \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing. We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data set.

@@

However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (here \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model. If this concerns an $s_2$ sample loaction, the associated response variable can no longer be used.

>>>>

To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 classes:

@@

As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample location, i.e. sample unit, does no longer consist of one, but multiple spatially agglomerated sample points. Whereas the maximal possible number of sample points per cluster $M$ is defined by the user, the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random. Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix). The package functions allow to define cluster-structures by specifying the argument \code{cluster}.

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{extended} and \textbf{pseudo extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured under this scenario.

@@

Comparing the \textit{extpsynth} and \textit{psmall} estimates, we see that in this particular case the point estimates are close and more important, the external as well as the design-based variances only differ marginally. This can be taken as evidence that the violation of the zero mean residual property can here expected to be negligible. We can also check the calculated mean of the residuals for the small area by

@@

As explained in section \ref{sec:str_and_mod}, a basis principle and prerequisite of the multi-phase estimation methods and estimators is that the sample phases ($s_0$, $s_1$, $s_2$) are \textit{nested} in each other. The correct nesting thereby concerns the spatial arrangement of the sample phases (figure \ref{fig:concmphase_and_sae_a}) as well as the availability of terrestrial and auxiliary information per phase and sample location. For the latter, the package \pkg{forestinventory} runs validity checks in the background, provides warning and error messages and, if possible, applies first aid adjustments to the inventory data set to prevent the calculations from failing. We will shortly demonstrate possible nesting violations by applying the gobal three-phase estimator to the \code{grisons} and \code{zberg} data sets.

@@

However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (here \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model. If this concerns an $s_2$ sample location, the associated response variable can no longer be used.
